---
title: "Banana peel weight"
subtitle: "A bunch of nonsense (mostly) about the banana data I collected"
author:
  - name: "Logan Groenestein"
    orcid: 0009-0008-2637-5923
    email: l.m.groenestein@uu.nl
date: 30 Sep 2026
date-format: "D MMM YYYY"
bibliography: exercise2bib.bib
format: 
  revealjs:
    logo: "crochetloon.jpeg"
    footer: Some footer text
    scrollable: true
---

```{r}
library(tidyverse)
library(plotly)
library(DT)
library(renv)
```

```{r}
#| cache: true

# Loading data and creating variables

data <- read_csv("Banana peel weight - Sheet1.csv") |> as.data.frame()

colnames(data) <- c("consumption", "purchase", "fullwgt", "peelwgt", 
                    "ediblewgt", "age")

data <- data[!is.na(data$consumption),]

# creating bunch variable based on purchase (not used here but still nice)
data$bunch <- factor(data$purchase)
levels(data$bunch) <- seq(1, length(levels(data$bunch)))
data$bunch <- as.numeric(data$bunch)

# centering variables of interest
data$fullwgt_c <- scale(data$fullwgt, scale = FALSE) |> as.numeric()
data$peelwgt_c <- scale(data$peelwgt, scale = FALSE) |> as.numeric()
data$ediblewgt_c <- scale(data$ediblewgt, scale = FALSE) |> as.numeric()
data$age_c <- scale(data$age, scale = FALSE) |> as.numeric()
```

## Research Question(s)

::: columns

::: {.column width="45%"}

Original question:

- Is the relationship between banana size and banana weight linear?

:::

::: {.column width="10%"}

:::

::: {.column width="45%"}

What I can (kind of) answer:

- Do banana peels get thinner (lighter) as the bananas age?

:::

:::

## My data

```{r}
#| label: tbl-data
#| tbl-cap: Banana data

data[,1:7] |> datatable(options = list(pageLength = 5))
```

Here in @tbl-data you can see the data I painstakingly collected for almost four months.

## Computing Centered Edible Weight

The (centered) "edible" weight of the bananas was not directly observed, so I had to compute it as


\begin{align*}
EdibleW_{C, i} & = EdibleW_i - \mu_{EdibleW} \\
& = FullW_i - PeelW_i - \mu_{EdibleW}, \\
\end{align*}


where $FullW_i$ is a banana's full weight including peel, $PeelW_i$ is the peel weight, and $\mu_{EdibleW}$ the mean "edible" banana weight, with $i = 1, \dots, n$. 

## Now in code

```{r}
#| echo: true
#| eval: false

data$ediblewgt <- data$fullwgt - data$peelwgt
data$ediblewgt_c <- scale(data$ediblewgt, scale = FALSE) |> as.numeric()
```

## Still figure

```{r}
#| label: fig-still
#| fig-cap: Plot of bananas' full weights and peel weights and age in days since purchase
ggplot(data, aes(x = fullwgt, y = peelwgt)) + 
  geom_point(aes(colour = age)) +
  xlab("Full weight of banana in grams") +
  ylab("Banana peel weight in grams") +
  theme_bw()
```

Here, @fig-still shows the relationship between bananas' full and peel weights, as well as their age.

## Interactive figure

```{r}
#| label: fig-wgt
#| fig-cap: Interactive plot of bananas' full weights and peel weights and age in days since purchase
p <- ggplot(data, aes(x = fullwgt, y = peelwgt)) + 
  geom_point(aes(colour = age)) +
  xlab("Full weight of banana in grams") +
  ylab("Banana peel weight in grams") +
  theme_bw()

ggplotly(p)
```

And @fig-wgt is an interactive version of @fig-still. 

## Citing a reference

I don't actually have any references for this, because I just kind of messed around by myself and did not look into any literature about bananas. Anyways, here's a source I had laying around in Zotero because I used it for my bachelor thesis @chambers.

## Reference(s)

::: {refs}
:::
